Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-08-27 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:08:27	16:38:57	(Thu27.Aug)

As will be demonstrated in upcoming sections, the LCA beamformer works very well with windows generated from the Kaiser-Bessel function. It is defined as:
%
\begin{align*}
\f_\beta = \bmat{
f_0(\beta) \\
\vdots\\
f_{M-1}(\beta)
}
\end{align*}
%
where
%
\begin{align}
f_m(\beta) = \frac{I_0\left(\pi\beta\sqrt{1-\left(\frac{2m}{M-1}-1\right)^2}\right)}{I_0(\pi\beta)}
\end{align}
%
and $I_0$ is the zeroth order modified Bessel function of the first kind:
%
\begin{align}
I_0(x) = \sumb{a=0}{\infty} \left[ \frac{\left(\frac{x}{2}\right)^a}{a!} \right]^2.
\end{align}
%
The Kaiser-Bessel window is near optimal in the sense of having its peak energy concentration around $\theta=0^\circ$, for a given time-bandwidth product specified with the parameter $\beta$:
%
\begin{align}
\beta = \frac{TB}{2},
\end{align}
%
where $T$ is the extent of the window in time and $B$ is its bandwidth. Adjusting $\beta$ changes the trade-off between mainlobe width and sidelobe level. At small values the windows approach the response of a rectangular window, and at large values ($>5$) the window converges to a Gaussian both in time and frequency. This class of windows is generally considered well suited for separating closely spaced sources with amplitudes of a high dynamic range~\cite{Harris1978}, they are easy to make, and they are optimal for any value of $\beta$.

@@

To allow LCA some added flexibility we also create variations of each window that are steered a little to each side.
%
\begin{align}
\w_{\beta,\phi} = \frac{\f_\beta\H\diag{(\a_\phi)}}{\f_\beta\H\a_\phi}\\
\end{align}
%
where
%
\begin{align}
\a_\phi = \bmat{
1 \\
e^{-j\frac{md}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{(M-1)d}{\lambda}\sin(\phi)}
}.
\end{align}
%
Here $d$ is the element spacing and $\lambda$ is the wavelength. The normalization factor $\f_\beta\H\a_\phi$ is the reciprocal of the window's coherent gain. It ensures unit gain in the direction of interest as required by (\ref{eq_lca_constraint}). Since the signal-to-noise ratio is constant for a specific window (same $\beta$-value), this normalization also proportionally increases the incoherent noise gain.

@@

There is an upper bound to the amount of steering we should apply to a window. Since each window's response is constrained to a unit gain in the look direction, each steered window's response is gained as shown in \Fig{windows_steering}. This increases the white noise gain and deteriorates the signal to noise ratio. To limit this we placing an upper bound to the steering, taking into account that a wide window can be allowed more steering than a narrow one.

We call the upper bound for steering $\phi_{\text{-3dB}}(\beta)$, and define it as the steering angle needed for the steered response to have a -3dB crossing that is exactly half that of the unsteered window. We performed a numeric search for the steering angle $\phi$ that fulfilled this criterion for a wide range of $\beta$-values. A third degree polynomial fit to these points is:
%
\begin{align}
\phi_\text{-3dB}(\beta) \approx -0.002894\beta^3 + 0.02451\beta^2 + -0.01308\beta + 0.2980
\end{align}
%
Any steering will be specified relative to this upper bound. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.

>>>>

As will be demonstrated in upcoming sections, the LCA beamformer works very well with windows generated from the Kaiser-Bessel function. Let it be expressed on vector form as:
%
\begin{align*}
\f_\beta = \bmat{
f_0(\beta) \\
\vdots\\
f_{M-1}(\beta)
}
\end{align*}
%
where
%
\begin{align}
f_m(\beta) = \frac{I_0\left(\pi\beta\sqrt{1-\left(\frac{2m}{M-1}-1\right)^2}\right)}{I_0(\pi\beta)}
\end{align}
%
and $I_0$ is the zeroth order modified Bessel function of the first kind:
%
\begin{align}
I_0(x) = \sumb{a=0}{\infty} \left[ \frac{\left(\frac{x}{2}\right)^a}{a!} \right]^2.
\end{align}
%
The Kaiser-Bessel window is near optimal in the sense of having its peak energy concentration around $\theta=0^\circ$, for a given time-bandwidth product specified with the parameter $\beta$:
%
\begin{align}
\beta = \frac{TB}{2},
\end{align}
%
where $T$ is the extent of the window in time and $B$ is its bandwidth. Adjusting $\beta$ changes the trade-off between mainlobe width and sidelobe level. When $\beta=0$ the window becomes a rectangular, while at large values ($\beta>5$) the window converges to a Gaussian both in time and frequency. This class of windows is generally considered well suited for separating closely spaced sources with amplitudes of a high dynamic range~\cite{Harris1978}, they are easy to make, and they are optimal for any value of $\beta$.

@@

To allow LCA some added flexibility we also create variations of each window that are steered a little to each side.
%
\begin{align}
\w_{\beta,\phi} = \frac{\f_\beta\H\diag{(\a_\phi)}}{\f_\beta\H\a_\phi}\\
\end{align}
%
where
%
\begin{align}
\a_\phi = \bmat{
1 \\
e^{-j\frac{md}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{(M-1)d}{\lambda}\sin(\phi)}
}.
\end{align}
%
Here $d$ is the element spacing, $\lambda$ is the wavelength and $\phi$ is the steering amount. The normalization factor $\f_\beta\H\a_\phi$ is the reciprocal of the window's coherent gain. It ensures unit gain in the direction of interest as required by (\ref{eq_lca_constraint}). Since the signal-to-noise ratio is constant for a specific window (same $\beta$-value), this normalization also proportionally increases the incoherent noise gain.

@@

There is an upper bound to the amount of steering we should apply to a window. Since each window's response is constrained to a unit gain in the look direction, each steered window's response is gained as shown in \Fig{windows_steering}. This increases the white noise gain and deteriorates the signal-to-noise ratio. To limit this we placing an upper bound to the steering, taking into account that a wide window can be allowed more steering than a narrow one.

We call the upper bound for steering $\phi_{\text{-3dB}}(\beta)$, and define it as the steering angle needed for the steered response to have a -3dB crossing that is exactly half that of the unsteered window. Using a numerical search for the steering angle $\phi$ that fulfilled this criterion for a wide range of $\beta$-values, we obtained a third degree polynomial fit to these points:
%
\begin{align}
\phi_\text{-3dB}(\beta) \approx -0.002894\beta^3 + 0.02451\beta^2 + -0.01308\beta + 0.2980
\end{align}
%
Any steering will be specified relative to this upper bound. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.
